# Experiment Description Data

# Experiment Values Data

The features selected for this database come from the accelerometer and gyroscope 3-axial raw signals calculated.time-XYZ and calculated.time.gyro-XYZ. These time domain signals (prefix 'time' to denote time) were captured at a constant rate of 50 Hz. Then they were filtered using a median filter and a 3rd order low pass Butterworth filter with a corner frequency of 20 Hz to remove noise. Similarly, the acceleration signal was then separated into body and gravity acceleration signals (calculated.time.body.Acc-XYZ and calculated.time.gravityAcc-XYZ) using another low pass Butterworth filter with a corner frequency of 0.3 Hz. 

Subsequently, the body linear acceleration and angular velocity were derived in time to obtain Jerk signals (calculated.time.body.AccJerk-XYZ and calculated.time.body.GyroJerk-XYZ). Also the magnitude of these three-dimensional signals were calculated using the Euclidean norm (calculated.time.body.AccMag, calculated.time.gravityAccMag, calculated.time.body.AccJerkMag, calculated.time.body.GyroMag, calculated.time.body.GyroJerkMag). 

Finally a Fast Fourier Transform (FFT) was applied to some of these signals producing calculated.frequency.body.Acc-XYZ, calculated.frequency.body.AccJerk-XYZ, calculated.frequency.body.Gyro-XYZ, calculated.frequency.body.AccJerkMag, calculated.frequency.body.GyroMag, calculated.frequency.body.GyroJerkMag. (Note the 'frequency' to indicate frequency domain signals). 

These signals were used to estimate variables of the feature vector for each pattern:  
'-XYZ' is used to denote 3-axial signals in the X, Y and Z directions.

## Define Constants

```{r const}
# Name of directory of the unzip data from url
dataset.directory <-'UCI HAR Dataset'

# Directory where the data should be downloaded
setwd( '~' )
dataset.download.dir <- getwd()
```

## Load Tidy File
```{r loadfile}
tidy.file       <- file.path( dataset.download.dir, 'tidy_data.txt')
classes <- rep( "numeric", 68 )
classes[1:2] <- "factor" 
tidy.data       <- read.table(
  tidy.file,
  stringsAsFactors=FALSE,
  header=TRUE,
  colClasses=classes
)
```

## overview
```{r overview_names}
names( tidy.data )
```
```{r overview_str}
str( tidy.data )
```
## subject

Id of the Person under the experiment

```{r subject_values}
unique(tidy.data$subject)
summary(tidy.data$subject)
```

## activity

Type of activity in pratice. 

Values:
```{r activity_values}
unique(tidy.data$activity)
summary(tidy.data$activity)
```

## Time Body accelerometer mean and standard deviation x-y-z

* calculated.time.bodyacc.mean.x
```{r summary_calculated_time_bodyacc_mean_x}
summary(tidy.data$calculated.time.bodyacc.mean.x)
```
* calculated.time.bodyacc.mean.y
```{r summary_calculated_time_bodyacc_mean_y}
summary(tidy.data$calculated.time.bodyacc.mean.y)
```
* calculated.time.bodyacc.mean.z
```{r summary_calculated_time_bodyacc_mean_z}
summary(tidy.data$calculated.time.bodyacc.mean.z)
```
* calculated.time.bodyacc.std.x
```{r summary_calculated_time_bodyacc_std_x}
summary(tidy.data$calculated.time.bodyacc.std.x)
```
* calculated.time.bodyacc.std.y
```{r summary_calculated_time_bodyacc_std_y}
summary(tidy.data$calculated.time.bodyacc.std.y)
```
* calculated.time.bodyacc.std.z
```{r summary_calculated_time_bodyacc_std_z}
summary(tidy.data$calculated.time.bodyacc.std.z)
```
## Frequence Body accelerometer mean and standard deviation x-y-z

* calculated.frequency.bodyacc.mean.x
```{r sumary_calculated.frequency.bodyacc.mean.x}
summary(tidy.data$calculated.frequency.bodyacc.mean.x)
```
* calculated.frequency.bodyacc.mean.y
```{r sumary_calculated.frequency.bodyacc.mean.y}
summary(tidy.data$calculated.frequency.bodyacc.mean.y)
```
* calculated.frequency.bodyacc.mean.z
```{r sumary_calculated.frequency.bodyacc.mean.z}
summary(tidy.data$calculated.frequency.bodyacc.mean.z)
```
* calculated.frequency.bodyacc.std.x
```{r sumary_calculated.frequency.bodyacc.std.x}
summary(tidy.data$calculated.frequency.bodyacc.std.x)
```
* calculated.frequency.bodyacc.std.y
```{r sumary_calculated.frequency.bodyacc.std.y}
summary(tidy.data$calculated.frequency.bodyacc.std.y)
```
* calculated.frequency.bodyacc.std.z
```{r sumary_calculated.frequency.bodyacc.std.z}
summary(tidy.data$calculated.frequency.bodyacc.std.z)
```

## Time Body gravity acceleration signals mean and standard deviation x-y-z

* calculated.time.gravityacc.mean.x
```{r_summary_calculated.time.gravityacc.mean.x}
summary(tidy.data$calculated.time.gravityacc.mean.x)
```
* calculated.time.gravityacc.mean.y
```{r_summary_calculated.time.gravityacc.mean.y}
summary(tidy.data$calculated.time.gravityacc.mean.y)
```
* calculated.time.gravityacc.mean.z
```{r_summary_calculated.time.gravityacc.mean.z}
summary(tidy.data$calculated.time.gravityacc.mean.z)
```
* calculated.time.gravityacc.std.x
```{r_summary_calculated.time.gravityacc.std.x}
summary(tidy.data$calculated.time.gravityacc.std.x)
```
* calculated.time.gravityacc.std.y
```{r_summary_calculated.time.gravityacc.std.y}
summary(tidy.data$calculated.time.gravityacc.std.y)
```
* calculated.time.gravityacc.std.z
```{r_summary_calculated.time.gravityacc.std.z}
summary(tidy.data$calculated.time.gravityacc.std.z)
```

## Time Body accelerometer mean and standard deviation of Jerk Signal x-y-z

* calculated.time.bodyaccjerk.mean.x
```{r summary_calculated.time.bodyaccjerk.mean.x}
summary(tidy.data$calculated.time.bodyaccjerk.mean.x)
```
* calculated.time.bodyaccjerk.mean.y
```{r summary_calculated.time.bodyaccjerk.mean.y}
summary(tidy.data$calculated.time.bodyaccjerk.mean.y)
```
* calculated.time.bodyaccjerk.mean.z
```{r summary_calculated.time.bodyaccjerk.mean.z}
summary(tidy.data$calculated.time.bodyaccjerk.mean.z)
```
* calculated.time.bodyaccjerk.std.x
```{r summary_calculated.time.bodyaccjerk.std.x}
summary(tidy.data$calculated.time.bodyaccjerk.std.x)
```
* calculated.time.bodyaccjerk.std.y
```{r summary_calculated.time.bodyaccjerk.std.y}
summary(tidy.data$calculated.time.bodyaccjerk.std.y)
```
* calculated.time.bodyaccjerk.std.z
```{r summary_calculated.time.bodyaccjerk.std.z}
summary(tidy.data$calculated.time.bodyaccjerk.std.z)
```

## Frequency Body accelerometer mean and standard deviation of Jerk Signal x-y-z

* calculated.frequency.bodyaccjerk.mean.x
```{r summary_calculated.frequency.bodyaccjerk.mean.x}
summary(tidy.data$calculated.frequency.bodyaccjerk.mean.x)
```
* calculated.frequency.bodyaccjerk.mean.y
```{r summary_calculated.frequency.bodyaccjerk.mean.y}
summary(tidy.data$calculated.frequency.bodyaccjerk.mean.y)
```
* calculated.frequency.bodyaccjerk.mean.z
```{r summary_calculated.frequency.bodyaccjerk.mean.z}
summary(tidy.data$calculated.frequency.bodyaccjerk.mean.z)
```
* calculated.frequency.bodyaccjerk.std.x
```{r summary_calculated.frequency.bodyaccjerk.std.x}
summary(tidy.data$calculated.frequency.bodyaccjerk.std.x)
```
* calculated.frequency.bodyaccjerk.std.y
```{r summary_calculated.frequency.bodyaccjerk.std.y}
summary(tidy.data$calculated.frequency.bodyaccjerk.std.y)
```
* calculated.frequency.bodyaccjerk.std.z
```{r summary_calculated.frequency.bodyaccjerk.std.z}
summary(tidy.data$calculated.frequency.bodyaccjerk.std.z)
```

## Time Body gyroscope mean and standard deviation x-y-z

* calculated.time.bodygyro.mean.x
```{r summary_calculated.time.bodygyro.mean.x}
summary(tidy.data$calculated.time.bodygyro.mean.x)
```
* calculated.time.bodygyro.mean.y
```{r summary_calculated.time.bodygyro.mean.y}
summary(tidy.data$calculated.time.bodygyro.mean.y)
```
* calculated.time.bodygyro.mean.z
```{r summary_calculated.time.bodygyro.mean.z}
summary(tidy.data$calculated.time.bodygyro.mean.z)
```
* calculated.time.bodygyro.std.x
```{r summary_calculated.time.bodygyro.std.x}
summary(tidy.data$calculated.time.bodygyro.std.x)
```
* calculated.time.bodygyro.std.y
```{r summary_calculated.time.bodygyro.std.y}
summary(tidy.data$calculated.time.bodygyro.std.y)
```
* calculated.time.bodygyro.std.z
```{r summary_calculated.time.bodygyro.std.z}
summary(tidy.data$calculated.time.bodygyro.std.z)
```

## Frequency Body gyroscope mean and standard deviation x-y-z

* calculated.frequency.bodygyro.mean.x
```{r summary_calculated.frequency.bodygyro.mean.x}
summary(tidy.data$calculated.frequency.bodygyro.mean.x)
```
* calculated.frequency.bodygyro.mean.y
```{r summary_calculated.frequency.bodygyro.mean.y}
summary(tidy.data$calculated.frequency.bodygyro.mean.y)
```
* calculated.frequency.bodygyro.mean.z
```{r summary_calculated.frequency.bodygyro.mean.z}
summary(tidy.data$calculated.frequency.bodygyro.mean.z)
```
* calculated.frequency.bodygyro.std.x
```{r summary_calculated.frequency.bodygyro.std.x}
summary(tidy.data$calculated.frequency.bodygyro.std.x)
```
* calculated.frequency.bodygyro.std.y
```{r summary_calculated.frequency.bodygyro.std.y}
summary(tidy.data$calculated.frequency.bodygyro.std.y)
```
* calculated.frequency.bodygyro.std.z
```{r summary_calculated.frequency.bodygyro.std.z}
summary(tidy.data$calculated.frequency.bodygyro.std.z)
```

## Time Body gyroscope mean and standard deviation of Jeark Signal x-y-z

* calculated.time.bodygyrojerk.mean.x
```{r summary_calculated.time.bodygyrojerk.mean.x}
summary(tidy.data$calculated.time.bodygyrojerk.mean.x)
```
* calculated.time.bodygyrojerk.mean.y
```{r summary_calculated.time.bodygyrojerk.mean.y}
summary(tidy.data$calculated.time.bodygyrojerk.mean.y)
```
* calculated.time.bodygyrojerk.mean.z
```{r summary_calculated.time.bodygyrojerk.mean.z}
summary(tidy.data$calculated.time.bodygyrojerk.mean.z)
```
* calculated.time.bodygyrojerk.std.x
```{r summary_calculated.time.bodygyrojerk.std.x}
summary(tidy.data$calculated.time.bodygyrojerk.std.x)
```
* calculated.time.bodygyrojerk.std.y
```{r summary_calculated.time.bodygyrojerk.std.y}
summary(tidy.data$calculated.time.bodygyrojerk.std.y)
```
* calculated.time.bodygyrojerk.std.z
```{r summary_calculated.time.bodygyrojerk.std.z}
summary(tidy.data$calculated.time.bodygyrojerk.std.z)
```

## Time Body accelerometer magnitude mean and standard deviation

* calculated.time.bodyaccmag.mean
```{r summary_calculated.time.bodyaccmag.mean}
summary(tidy.data$calculated.time.bodyaccmag.mean)
```
* calculated.time.bodyaccmag.std
```{r summary_calculated.time.bodyaccmag.std}
summary(tidy.data$calculated.time.bodyaccmag.std)
```

## Frequency Body accelerometer magnitude mean and standard deviation

* calculated.frequency.bodyaccmag.mean
```{r summary_calculated.frequency.bodyaccmag.mean}
summary(tidy.data$calculated.frequency.bodyaccmag.mean)
```
* calculated.frequency.bodyaccmag.std
```{r summary_calculated.frequency.bodyaccmag.std}
summary(tidy.data$calculated.frequency.bodyaccmag.std)
```


## Time Gravity accelerometer magnitude mean and standard deviation

* calculated.time.gravityaccmag.mean
```{r summary_calculated.time.gravityaccmag.mean}
summary(tidy.data$calculated.time.gravityaccmag.mean)
```
* calculated.time.gravityaccmag.std
```{r summary_calculated.time.gravityaccmag.std}
summary(tidy.data$calculated.time.gravityaccmag.std)
```

## Time Body acceleromenter jerk magnitude mean and standard deviation

* calculated.time.bodyaccjerkmag.mean
```{r summary_calculated.time.bodyaccjerkmag.mean}
summary(tidy.data$calculated.time.bodyaccjerkmag.mean)
```
* calculated.time.bodyaccjerkmag.std
```{r summary_calculated.time.bodyaccjerkmag.std}
summary(tidy.data$calculated.time.bodyaccjerkmag.std)
```

## Frequency Body acceleromenter jerk magnitude mean and standard deviation

* calculated.frequency.bodyaccjerkmag.mean
```{r summary_calculated.frequency.bodyaccjerkmag.mean}
summary(tidy.data$calculated.frequency.bodyaccjerkmag.mean)
```
* calculated.frequency.bodyaccjerkmag.std
```{r summary_calculated.frequency.bodyaccjerkmag.std}
summary(tidy.data$calculated.frequency.bodyaccjerkmag.std)
```

## Time Body of gyroscope magnitude mean and standard deviation

* calculated.time.bodygyromag.mean
```{r summary_calculated.time.bodygyromag.mean}
summary(tidy.data$calculated.time.bodygyromag.mean)
```
* calculated.time.bodygyromag.std
```{r summary_calculated.time.bodygyromag.std}
summary(tidy.data$calculated.time.bodygyromag.std)
```

## Frequency Body of gyroscope magnitude mean and standard deviation

* calculated.frequency.bodygyromag.mean
```{r summary_calculated.frequency.bodygyromag.mean}
summary(tidy.data$calculated.frequency.bodygyromag.mean)
```
* calculated.frequency.bodygyromag.std
```{r summary_calculated.frequency.bodygyromag.std}
summary(tidy.data$calculated.frequency.bodygyromag.std)
```

## Time Body of gyroscope jerk magnitude mean and standard deviation

* calculated.time.bodygyrojerkmag.mean
```{r summary_calculated.time.bodygyrojerkmag.mean}
summary(tidy.data$calculated.time.bodygyrojerkmag.mean)
```
* calculated.time.bodygyrojerkmag.std
```{r summary_calculated.time.bodygyrojerkmag.std}
summary(tidy.data$calculated.time.bodygyrojerkmag.std)
```

## Frequency Body of gyroscope jerk magnitude mean and standard deviation

* calculated.frequency.bodygyrojerkmag.mean
```{r summary_calculated.frequency.bodygyrojerkmag.mean}
summary(tidy.data$calculated.frequency.bodygyrojerkmag.mean)
```
* calculated.frequency.bodygyrojerkmag.std
```{r summary_calculated.frequency.bodygyrojerkmag.std}
summary(tidy.data$calculated.frequency.bodygyrojerkmag.std)
```

The set of variables that were estimated from these signals are: 

```
mean(): Mean value
std(): Standard deviation
mad(): Median absolute deviation 
max(): Largest value in array
min(): Smallest value in array
sma(): Signal magnitude area
energy(): Energy measure. Sum of the squares divided by the number of values. 
iqr(): Interquartile range 
entropy(): Signal entropy
arCoeff(): Autorregresion coefficients with Burg order equal to 4
correlation(): correlation coefficient between two signals
maxInds(): index of the frequency component with largest magnitude
meanFreq(): Weighted average of the frequency components to obtain a mean frequency
skewness(): skewness of the frequency domain signal 
kurtosis(): kurtosis of the frequency domain signal 
bandsEnergy(): Energy of a frequency interval within the 64 bins of the FFT of each window.
angle(): Angle between to vectors.
```
